`State and Parameter Estimation for Spatio-Temporal Neural Fields' or `A Data Driven Framework for Patient-Specific Neural Field Modelling'

### 2010-06-22 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
some more general edits

\section{Discussion}\label{DiscussionSection}
This paper has presented a method for estimating intracortical connectivity and synaptic dynamics from electrophysiological data. The results indicate good estimation performance, where the mean parameter estimates are in good agreement with the actual values. The estimates of the connectivity kernel parameters show relatively broad distributions, which is expected considering the system is driven by an unknown input, that is modelled as a random disturbance, and the observations are corrupted by noise. To overcome the effect of noise, an evoked-response paradigm should be considered when applying the estimation algorithm to electrophysiological recordings. Following such a paradigm, data can be epoched into a sufficient number of trials (realisations), where each trial is a response to a stimulus, such that a distribution of the parameter estimates can be formed.

A significant contribution of this paper is the derivation of the basis function decomposition of the neural field model, as described in Section~\ref{Sect:ReducedModelDerivation}. Through the derivation, a finite-dimensional state-space model is formed that approximates a continuous neural field. Previous work from Schiff and Sauer~\cite{schiff2008kalman} demonstrated for the first time the applicability of the unscented Kalman filter for state estimation for discrete (in space and time) nonlinear spatio-temporal systems. Within their framework, each discrete spatial location forms an element of the state vector to be estimated with the UKF. A limitation of this framework lies in an increase in dimensionality with increases in  the spatial resolution and in the physical size of the neural field. Increases in dimensionality lead to increases in uncertainty of the estimated state and problems associated with computational complexity. This paper builds on this previous study, where the estimator is constructed to approximate a continuous field and the dimensionality of the state vector is independent of the resolution of a spatial discretisation and is governed by the inherent spatial properties of the field. The frequency analysis methods presented in Section~\ref{SpectralAnalysisSection} provide a rigorous framework for the placement and the number of the basis functions required for the decomposition. This enables estimation of a larger field, with a lower dimensional state vector. The basis function decomposition also enabled the coupling to be described by an inner product rather than a spatial convolution. This transformation greatly simplifies the computational complexity of the estimator, which is important for the estimator to run in a real-time scenario.

@@

State disturbance defined in equation~\ref{eq:Wt} is a linear function of $e_t(\mathbf r)$ and hence $\mathbf{e}_t$ is also normally distributed. The expected value of $\mathbf e_t$ is given by 

@@

	&=&\mathbf 0 

@@

	&=&\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}(\mathbf r) \gamma(\mathbf r- \mathbf r' )\boldsymbol{\phi}^{\top}(\mathbf r')d\mathbf r' d\mathbf r\mathbf{\Gamma}^{-\top} 

@@

&\times& \mathrm{exp}\left(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu\right),

@@

@@

   \boldsymbol\Phi(\boldsymbol\nu)=(\pi\sigma_{\phi}^2)^{\frac{n}{2}}\mathrm{exp}(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu).

@@

Substituting $\sigma_{\nu}^2=\frac{1}{\pi^2\sigma_{\phi}^2}$ gives 

@@

\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right).

@@

For $3$~dB attenuation at $\boldsymbol\nu_c$ we need to set

>>>>

\section{Discussion}\label{DiscussionSection}
This paper has presented a method for estimating intracortical connectivity and synaptic dynamics from electrophysiological data. As seen in the Section~\ref{sec:state_and_param_results} the results indicate good estimation performance, where the mean parameter estimates are in good agreement with the actual values. The estimates of the connectivity kernel parameters show relatively broad distributions, which is expected considering the system is driven by an unknown input, modelled as a random disturbance, and the observations are corrupted by noise. To overcome the effect of noise, an evoked-response paradigm should be considered when applying the estimation algorithm to electrophysiological recordings. Following such a paradigm, data can be epoched into a sufficient number of trials (realisations), where each trial is a response to a stimulus, such that a distribution of the parameter estimates can be formed.

Previous work from Schiff and Sauer~\cite{schiff2008kalman} demonstrated for the first time the applicability of the unscented Kalman filter (UKF) for state estimation for discrete (in space and time) nonlinear spatio-temporal systems. Within their framework, each discrete spatial location forms an element of the state vector to be estimated with the UKF. A limitation of this framework lies in an increase in dimensionality with increases in the spatial resolution and in the physical size of the neural field. Increases in dimensionality lead to increases in uncertainty of the estimated state and problems associated with computational complexity. This paper builds on this previous study, where the estimator is constructed to approximate a continuous field and the dimensionality of the state vector is independent of the resolution of a spatial discretisation and is governed by the inherent spatial properties of the field. The frequency analysis methods presented in Section~\ref{SpectralAnalysisSection} provide a rigorous framework for the placement and the number of the basis functions required for the decomposition. This enables estimation of a larger field, with a lower dimensional state vector. The basis function decomposition also enabled the coupling to be described by an inner product rather than a spatial convolution. This transformation greatly simplifies the computational complexity of the estimator, which is important for the estimator to run in a real-time scenario.

@@

The state disturbance defined in equation~\ref{eq:Wt} is a linear function of $e_t(\mathbf r)$, therefore $\mathbf{e}_t$ is also normally distributed. The expected value of $\mathbf e_t$ is given by 

@@

	&=&\mathbf 0.

@@

	&=&\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}(\mathbf r) \gamma(\mathbf r- \mathbf r' )\boldsymbol{\phi}^{\top}(\mathbf r')d\mathbf r' d\mathbf r\mathbf{\Gamma}^{-\top}.

@@

&\times& \mathrm{exp}\left(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu\right).

@@

Now evaluating the integral on the RHS of equation~\ref{eq:AppendixGaussianFT} gives 

@@

\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),

@@

% 
%    \boldsymbol\Phi(\boldsymbol\nu)=(\pi\sigma_{\phi}^2)^{\frac{n}{2}}\mathrm{exp}(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu).

@@

	\sigma_{\nu}^2=\frac{1}{\pi^2\sigma_{\phi}^2}. 

@@

For $3$~dB attenuation ($1/2$ to peak power) at the cut-off frequency, $\boldsymbol\nu_c$, we need to set
